---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# purrr
# stringr
# lubridate
#dplyr
# ggdist
# gghalves
```



```{r}
# look at control files to read in run parameters
# run_models <- read.csv("../data_cache/Controls/run_models.csv")[, "run_models"]



```

[] Check if run_models is TRUE
[x] Find the most recent file
[x] Find the second most recent file
[] Are there changes?
  [] Yes --
    [] Remake report
      [] QC
      [] Vis
      [] Stats
    [] Highlight new data
    [] Add a trigger to email out the new document in ../data_cache/Controls
  [] No -- 
    [] Alternate no changes document or send the old one with no changes noted.



```{r eval=FALSE, include=FALSE}
records <- list.files("./inst/extdata/")
records 
# 
# # extract creation times for each file from its name.
# records_times <- purrr::map(
#   records, function(record){
#     if ( stringr::str_detect(record, 'rootbot_') ){
#       record <- stringr::str_split(record, '_')
#       record <- stringr::str_remove(record, 'rootbot_')
#       return(lubridate::ymd_hms(record))  
#     } else {
#       # if there is not rootbot prefix, then we don't want to consider the file. 
#       # it must have been introduced by mistake. 
#       return(lubridate::ymd_hms("1970-01-01_00?00?00"))    
#     }  
# })
# 
# # consider the most recent two entries.
# newest_time        <- max(unlist(records_times))
# second_newest_time <- max(unlist(records_times[records_times != newest_time]))
# 
# 
# new_record <- read.csv(paste0("../data_cache/Records/", 
#                               records[records_times == newest_time]))
# 
# old_record <- read.csv(paste0("../data_cache/Records/", 
#                               records[records_times == second_newest_time]))
# # drop index
# new_record <- new_record[, !(names(new_record) %in% c('X'))]
# old_record <- old_record[, !(names(old_record) %in% c('X'))]
# 
# # Are there changes in the data?
# old_record['InOld'] <- TRUE 
# all_records <- dplyr::full_join(new_record, old_record)
# all_records[is.na(all_records$InOld), 'InOld'] <- FALSE
# 
# 
# # added records
# compare_records <- all_records[all_records$InOld == FALSE, ]
# 
# 
# n_new_records <- nrow(compare_records)
# if (n_new_records == 0){
#   print("No additional records found.")
#   # TODO write out to /Controls
#   # TODO take no-update action
# } else {
#   print(paste0(as.character(n_new_records)," new records found!"))
# }
# 
# 
# # TODO use the groups in compare_records to retrieve the relevant matches from 
# # all_records
# 
# head(compare_records)
# 
# # genotype
# # group
# # TODO tlc on group, some don't have a value.
```



```{r}
library(tidyverse)
# cleaning & QC
library(janitor)
```

## Load in metadata
The referenced files match naming of the xlsx file in teams. To have plain text files rather than binary the sheets are stored as individual csvs in `seeds in rootbot/`.
```{r}
# if(!file.exists("./data/metadata.rds")){
#TODO after drafting switch to above line
if(TRUE){
  metadata <- read.csv("./inst/extdata/seeds in rootbot/Log.csv")
  # remove human readable but not computationally practical columns of text.
  metadata <- metadata[, c("ï..282.inbred", 
                           "Rootbot.date", 
                           "scored", 
                           "source...200k.packets.from", 
                           "WW.cassette.", 
                           "WS.cassette.")]
  
  metadata <- metadata %>% 
    rename("inbred" = "ï..282.inbred",
           "rootbot_date" = "Rootbot.date", 
           "scored" = "scored", 
           "source" = "source...200k.packets.from", 
           "ww_cassette" = "WW.cassette.", 
           "ws_cassette" = "WS.cassette.") 
  
  saveRDS(metadata, file = "./data/metadata.rds")
}else{
  # metadata <- readRDS(file = "./data/metadata.rds")
}
```
## Load in records
```{r}
# if the aggregated dataframe (df.rds) exists load it otherwise build and save it
if(!file.exists("./data/df.rds")){
#TODO after drafting switch to above line
# if(TRUE){
  records <- list.files("./inst/extdata/plate_csvs")
  
  record_list <- list()
  for(record in records){
    # print(record) # for debugging
    entry <- read.csv(paste("./inst/extdata/plate_csvs", record, sep = "/"))
    entry["file"] <- record
    
    record_list[[length(record_list)+1]] <- entry
  }
  
  # sloppy and inefficent but functional.
  agg_df <- record_list[[1]]
  for(i in seq(2, length(record_list))){
    agg_df <- full_join(agg_df, record_list[[i]])
  }
  
  df <- janitor::clean_names(agg_df)
  
  
  
  # fix misnamed values
  root_name_list <- list(c(" B73", " B73seed#1"),
                         c(" C123_seed1", " C123seed#1")
                         )
  
  for(e in root_name_list){
    df[df$root_name == e[1], "root_name"] <- e[2]
  }
  
  df <- df[df$root_name != "", ]
  
  # split root_name into genotype/rep
  df <- separate(
    df,
    root_name,
    c("genotype", "replicate"),
    sep = "seed#",
    remove = TRUE,
    convert = FALSE,
    extra = "warn",
    fill = "warn"
  )
  
  # strip out leading space
  df$genotype <- str_remove_all(df$genotype, " ")
  

  
  saveRDS(df, file = "./data/df.rds")  
  
}else{
  df <- readRDS(file = "./data/df.rds")
}
```

```{r}
df <- df[, c(
  "length",        # y
  "genotype",      # Genotype
  # "time_elapsed",  # Time
  # "group",         # Well watered or Water Stressed  
  "replicate",     # seed within UID  
  # "uid",           # UID of plate x batch
  "image"
)]
```


## tidy up metadata
```{r}
metadata <- metadata %>% as_tibble()

metadata <- metadata %>% 
  # ignore non scored entries
  mutate(
    scored = case_when(
      scored == "yes" ~ TRUE,
      scored == "yes " ~ TRUE,
      scored == "no" ~ FALSE,
      scored == ""~ FALSE)
    ) %>% 
  filter(scored == TRUE)
  

metadata <- metadata %>% 
  # drop missing dates
  filter(!(rootbot_date %in% c("", " "))) %>% 
  # make into dates
  mutate(rootbot_date = lubridate::mdy(rootbot_date)) 
  
  
# Make tidy wrt cassette
metadata <- metadata %>% 
  pivot_longer(c("ww_cassette", "ws_cassette")) %>% 
  rename("condition" = "name",
         "cassette" = "value") %>% 
  mutate(condition = case_when(
    condition == "ww_cassette" ~ "ww",
    condition == "ws_cassette" ~ "ws"
  ))


metadata <- metadata %>% distinct()

metadata <- metadata %>% 
  mutate(uid = seq(1, nrow(metadata)))

```





## Tidy up data
```{r}
df <- as_tibble(df)

df["image_date"] <- stringr::str_extract(df$image, "\\d+") %>% lubridate::ymd()

df["image_datetime"] <- stringr::str_extract(df$image, "\\d+-\\d+") %>% lubridate::ymd_hms()

df["image_cassette"] <- stringr::str_extract(df$image, "\\d+$") %>% as.integer()


df["uid"] <- NA

# must contain non-missing values for these entries:
df <- df[(
  (!is.na(df$genotype)) & 
  (!is.na(df$image_cassette)) & 
  (!is.na(df$image_date))),]
```



```{r}
# uid_code <- 303 # 

for(uid_code in unique(metadata$uid)){
  exp_days_threshold = "4 days"
  # print(uid_code)
  match_inbred   <- metadata[metadata$uid == uid_code, "inbred"][[1]]
  match_cassette <- metadata[metadata$uid == uid_code, "cassette"][[1]]
  fuzzy_date     <- metadata[metadata$uid == uid_code, "rootbot_date"][[1]] # select first index because it's a tibble

  mask <- (df$genotype == match_inbred
      ) & (df$image_cassette == match_cassette
      ) & (difftime(fuzzy_date,
                    df$image_date,
                    units = 'days') < exp_days_threshold
      ) & (!is.na(df$genotype))
  
  # if there are matches
  if(mean(mask) > 0){
    df[mask, "uid"] <- uid_code
  }
}


# now use the uid as a grouping to find the time relative to first image
df <- df %>% 
  group_by(uid) %>% 
  mutate(min_datetime = min(image_datetime)) %>% 
  mutate(delta_datetime = image_datetime - min_datetime)


df
```


## merge
```{r}
data <- left_join(metadata, df) %>% filter(!is.na(length))

data 
```



```{r}
data %>% 
  filter(inbred == "C123") %>% 
  mutate(delta_datetime = delta_datetime/(60*60)) %>% # rescale from seconds to hours
  ggplot(aes(delta_datetime, length, color = condition))+
  geom_point()+
  geom_smooth(method = "lm", aes(fill = condition), alpha = 0.3)+
  # facet_grid(.~condition)
  labs(x = "Hours Elapsed", y = "Root Length (mm)", title = "Genotype: C123")
  
```


```{r}
# where ws is longer than ww
# slope or abs?\
library(broom)


M <- data %>% 
  filter(inbred == "C123") %>% 
  mutate(delta_datetime = delta_datetime/(60*60)) 


fm <- lm("length ~ delta_datetime*condition", M)
res <- broom::tidy(fm)

# is there a significant difference?
res[res$term == "delta_datetime:conditionww", "p.value"] < 0.05

# ws has a more positive slope than ww
res[res$term == "delta_datetime:conditionww", "estimate"] < 0



# find all inbreds without both conditions:
rm_geno <- data %>% 
  group_by(inbred) %>% 
  select(condition) %>% 
  distinct() %>% 
  tally() %>% 
  filter(n <2)
  
data <- data[!(data$inbred %in% rm_geno$inbred), ]




res_list <- purrr::map(unique(data$inbred), function(e){
  M <- data %>% 
  filter(inbred == e) %>% 
  mutate(delta_datetime = delta_datetime/(60*60)) 

  fm <- lm("length ~ delta_datetime*condition", M)
  res <- broom::tidy(fm)
  
})

res_list_sig <- purrr::map(res_list, function(res){
  # is there a significant difference?
  unlist(res[res$term == "delta_datetime:conditionww", "p.value"] )
})


res_list_est <- purrr::map(res_list, function(res){
  # ws has a more positive slope than ww
  unlist(res[res$term == "delta_datetime:conditionww", "estimate"])
})

res_overview <- data.frame(
  list_idx = seq(1, length(res_list)),
  p_value = unlist(res_list_sig),
  interact = unlist(res_list_est)
)
```

```{r}
res_overview %>%
  filter(!is.na(p_value)) %>% 
  # filter(p_value < 0.05) %>% 
  group_by(p_value < 0.05, interact < 0) %>% 
  tally()
```


# visualize those with ws favored
```{r}
list_idxs <- res_overview %>% filter(#p_value < 0.05, 
                                     interact < 0) %>% select(list_idx) %>% unlist()

i = 1
res_list[[list_idxs[i]]]


unique(data$inbred)[list_idxs[i]]


plts <- purrr::map(#list_idxs,
  c(15, #17, 18, 
    22, 32, 54, 
    #83, 
    99, 106),
                   function(ii){
  geno <- unique(data$inbred)[ii]
  
  data %>% 
  filter(inbred == geno) %>% 
  mutate(delta_datetime = delta_datetime/(60*60)) %>% # rescale from seconds to hours
  mutate(condition = factor(condition, levels = c("ww", "ws"))) %>% 
  ggplot(aes(delta_datetime, length, color = condition))+
  geom_point()+
  geom_smooth(method = "lm", aes(fill = condition), alpha = 0.3)+
  # facet_grid(.~condition)
  labs(x = "Hours Elapsed", y = "Root Length (cm)", title = paste0("Genotype: ", geno))+
  theme_bw()+
  theme(legend.position = "bottom")+
  ggsci::scale_color_lancet()+
  ggsci::scale_fill_lancet()
})

library(cowplot)
cowplot::plot_grid(plotlist = plts)

# end 11/14/2022

```





```{r}
geno <- "CML91"



data %>% 
  filter(inbred == geno) %>% 
  mutate(delta_datetime = delta_datetime/(60*60)) %>% # rescale from seconds to hours
  ggplot(aes(delta_datetime, length, color = condition))+
  geom_point()+
  geom_smooth(method = "lm", aes(fill = condition), alpha = 0.3)+
  # facet_grid(.~condition)
  labs(x = "Hours Elapsed", y = "Root Length (mm)", title = paste0("Genotype: ", geno))



```







```{r}


library(ggh4x)
library(ggsignif)  
library("PNWColors")
library('agricolae')

current_genotype = 'CML277'


for(current_genotype in unique(all_records$genotype)){
  
  temp <- all_records[all_records$genotype == current_genotype, ]
  
  # tweak group ordering for plotting
  temp$group <- factor(temp$group, c('ww', 'ws'))
  temp
  

  

  if( length(unique(temp$hours_elapsed)) == 1 ){

    
    fm_ph <- agricolae::HSD.test(fm, trt = c('group'))
    fm_ph <- fm_ph$groups
    fm_ph$comparison <- rownames(fm_ph)
    fm_ph$comparison <- stringr::str_replace(fm_ph$comparison, ":", ".")
    
  } else if( length(unique(temp$group)) == 1 ){

    
    fm_ph <- agricolae::HSD.test(fm, trt = c('hours_elapsed'))
    fm_ph <- fm_ph$groups
    fm_ph$comparison <- rownames(fm_ph)
    fm_ph$comparison <- stringr::str_replace(fm_ph$comparison, ":", ".")
    
  } else{
    fm <- lm(length ~ hours_elapsed*group, temp)
    # annotations
    fm_tab <- broom::tidy(fm)
    fm_tab    
    
    fm_ph <- agricolae::HSD.test(fm, trt = c('hours_elapsed', 'group'))
    fm_ph <- fm_ph$groups
    fm_ph$comparison <- rownames(fm_ph)
    fm_ph$comparison <- stringr::str_replace(fm_ph$comparison, ":", ".")
  }
  

  
  
  # set up base of the figure. Pass in aes to each geom so that we can use 
  # different columns for annotation
  fig <- temp %>% 
    ggplot()+
    ggdist::stat_halfeye(
      aes(interaction(hours_elapsed, group), length, fill = group),
      adjust = .5, 
      width = .6, 
      .width = 0, 
      justification = -.2, 
      point_colour = NA
    ) + 
    geom_boxplot(
      aes(interaction(hours_elapsed, group), length, fill = group),
      width = .15, 
      outlier.shape = NA
    ) +
    ## add justified jitter from the {gghalves} package
    gghalves::geom_half_point(
          aes(interaction(hours_elapsed, group), length, fill = group),
          side = "l", 
      range_scale = .4, 
      alpha = .3
    ) +
    scale_x_discrete(guide = "axis_nested")+
    xlab("")+
    ylab("Length (mm)")+
    labs(title = paste('Genotype: ', current_genotype))+
    ggthemes::theme_clean()+
    ggthemes::scale_fill_few()+
    theme(legend.position = 'bottom')
  
  
  
  # check if there's more than one group and add in the relevant comparisons 
  if( !((length(unique(temp$group)) == 1) & (length(unique(temp$hours_elapsed)) == 1)) ){

    y_pos_start <- max(temp$length)
    y_pos_step <- 1 #y_pos_start / 10
    # select each hsd group. Then for each letter in in the grouping code add in a 
    # label with a letter specific y axis
    for(i in seq_along(unique(fm_ph$comparison)) ){
      fm_ph_subset = fm_ph[fm_ph$comparison == unique(fm_ph$comparison)[i] , ]
      
      # this is a messy way to get the comparisons at different locations.
      hsd_letters <- strsplit(fm_ph_subset$groups, split = "")[[1]]
      for(hsd_letter in hsd_letters){
        y_increment = seq_along(letters)[hsd_letter == letters]
        
        dat = data.frame(
            x=fm_ph_subset$comparison,
            y=c(y_pos_start+(y_pos_step*(y_increment))),
            annotation=toupper(fm_ph_subset$groups))
        
        fig <- fig+geom_label(
          data=dat,
          aes(x=x, y=y, label=annotation), 
          fontface="bold", 
          fill = pnw_palette("Sailboat",4,type="continuous")[y_increment],
          color = pnw_palette("Sailboat",4,type="continuous")[y_increment]
          )
        
        fig <- fig+geom_text(
          data=dat,
          aes(x=x, y=y, label=annotation), 
          fontface="bold"
          )
      }
    }    
    
        
  }
  
  
  

  
  
  ggsave(paste0('../reports/figures/',current_genotype,'.tiff'), fig)

}





```





























```{r}
# Need to match up images by date
# Need to match up plates to condition


# Start with metadata, query data for matching entries.
  # use an indicator column for which obs should be dropped.
  # because metadata:obs is one to many metadata will be merged into data

# 
```






```{r}


```






```{r}

# 
all_records <- read.csv('../data/cache/rootbot_2022-03-03_14-20-06.464849.csv')
```


```{r}
library(tidyverse)


# cleaning & QC
library(janitor)
all_records <- janitor::clean_names(all_records)


all_records <- all_records[, c(
  "length",        # y
  "time_elapsed",  # Time
  "group",         # Well watered or Water Stressed
  "genotype",      # Genotype
  "uid",           # UID of plate x batch
  "replicate"      # seed within UID
  )]

```



```{r}
# Round down elapsed time into something more useful


all_records$days_elapsed <- str_extract(all_records$time_elapsed, "\\d days") %>%
  str_replace(' days', '') %>% as.numeric()
all_records$sec_elapsed <- str_extract(all_records$time_elapsed, "\\d+:\\d+:\\d+") %>% 
  lubridate::hms() %>% 
  lubridate::as.difftime()

# Add days elapsed in after converting to seconds.
all_records$sec_elapsed <- all_records$sec_elapsed + (all_records$days_elapsed*(24*60*60))
all_records$sec_elapsed <- lubridate::as.difftime(all_records$sec_elapsed)


all_records$hours_elapsed <- round(all_records$sec_elapsed / (60*60), 0) %>% as.integer()

all_records <- all_records[, c("length", "hours_elapsed", "group", "genotype", "uid", "replicate")]


# summary statistics
ggplot(all_records, aes(hours_elapsed))+
  geom_histogram(bins =  100)

#   geom_density()



# TODO there are some values which do not fall in the expected two times. 
# These should be handled by 
# 1. allowing factor levels to vary between groups or 
# 2. widening the bins (e.g. 27-33 -> "30")


# summary statistics
tally_tab <- all_records %>% 
  select(genotype, group, length) %>% 
  group_by(genotype, group) %>% 
  tally()


summary_tab <- all_records %>% 
  select(genotype, group, length) %>% 
  group_by(genotype, group) %>% 
  summarise(mean_length = mean(length, na.rm = T),
            sd_length   = sd(length, na.rm = T)
            )

summary_tab


```

```{r}
library(ggh4x)
library(ggsignif)  
library("PNWColors")
library('agricolae')

current_genotype = 'CML277'


for(current_genotype in unique(all_records$genotype)){
  
  temp <- all_records[all_records$genotype == current_genotype, ]
  
  # tweak group ordering for plotting
  temp$group <- factor(temp$group, c('ww', 'ws'))
  temp
  

  

  if( length(unique(temp$hours_elapsed)) == 1 ){

    
    fm_ph <- agricolae::HSD.test(fm, trt = c('group'))
    fm_ph <- fm_ph$groups
    fm_ph$comparison <- rownames(fm_ph)
    fm_ph$comparison <- stringr::str_replace(fm_ph$comparison, ":", ".")
    
  } else if( length(unique(temp$group)) == 1 ){

    
    fm_ph <- agricolae::HSD.test(fm, trt = c('hours_elapsed'))
    fm_ph <- fm_ph$groups
    fm_ph$comparison <- rownames(fm_ph)
    fm_ph$comparison <- stringr::str_replace(fm_ph$comparison, ":", ".")
    
  } else{
    fm <- lm(length ~ hours_elapsed*group, temp)
    # annotations
    fm_tab <- broom::tidy(fm)
    fm_tab    
    
    fm_ph <- agricolae::HSD.test(fm, trt = c('hours_elapsed', 'group'))
    fm_ph <- fm_ph$groups
    fm_ph$comparison <- rownames(fm_ph)
    fm_ph$comparison <- stringr::str_replace(fm_ph$comparison, ":", ".")
  }
  

  
  
  # set up base of the figure. Pass in aes to each geom so that we can use 
  # different columns for annotation
  fig <- temp %>% 
    ggplot()+
    ggdist::stat_halfeye(
      aes(interaction(hours_elapsed, group), length, fill = group),
      adjust = .5, 
      width = .6, 
      .width = 0, 
      justification = -.2, 
      point_colour = NA
    ) + 
    geom_boxplot(
      aes(interaction(hours_elapsed, group), length, fill = group),
      width = .15, 
      outlier.shape = NA
    ) +
    ## add justified jitter from the {gghalves} package
    gghalves::geom_half_point(
          aes(interaction(hours_elapsed, group), length, fill = group),
          side = "l", 
      range_scale = .4, 
      alpha = .3
    ) +
    scale_x_discrete(guide = "axis_nested")+
    xlab("")+
    ylab("Length (mm)")+
    labs(title = paste('Genotype: ', current_genotype))+
    ggthemes::theme_clean()+
    ggthemes::scale_fill_few()+
    theme(legend.position = 'bottom')
  
  
  
  # check if there's more than one group and add in the relevant comparisons 
  if( !((length(unique(temp$group)) == 1) & (length(unique(temp$hours_elapsed)) == 1)) ){

    y_pos_start <- max(temp$length)
    y_pos_step <- 1 #y_pos_start / 10
    # select each hsd group. Then for each letter in in the grouping code add in a 
    # label with a letter specific y axis
    for(i in seq_along(unique(fm_ph$comparison)) ){
      fm_ph_subset = fm_ph[fm_ph$comparison == unique(fm_ph$comparison)[i] , ]
      
      # this is a messy way to get the comparisons at different locations.
      hsd_letters <- strsplit(fm_ph_subset$groups, split = "")[[1]]
      for(hsd_letter in hsd_letters){
        y_increment = seq_along(letters)[hsd_letter == letters]
        
        dat = data.frame(
            x=fm_ph_subset$comparison,
            y=c(y_pos_start+(y_pos_step*(y_increment))),
            annotation=toupper(fm_ph_subset$groups))
        
        fig <- fig+geom_label(
          data=dat,
          aes(x=x, y=y, label=annotation), 
          fontface="bold", 
          fill = pnw_palette("Sailboat",4,type="continuous")[y_increment],
          color = pnw_palette("Sailboat",4,type="continuous")[y_increment]
          )
        
        fig <- fig+geom_text(
          data=dat,
          aes(x=x, y=y, label=annotation), 
          fontface="bold"
          )
      }
    }    
    
        
  }
  
  
  

  
  
  ggsave(paste0('../reports/figures/',current_genotype,'.tiff'), fig)

}









```



```{r}

```







```{r}

```


```{r}

temp = all_records[all_records$genotype == ' NC338', ]


lm(length ~ group, temp)


```


```{r}
skimr::skim(all_records)
```


```{r}
library(lme4)

lmer(length ~ genotype*group , random = ~1|rootid)

```

